Lab 12: Covariate Balancing and CBPS

### 2023-04-12 Lab 13  CBPS/Lab12_covariatebalance.Rmd
Add files via upload

---
title: "Lab 13: Covariate Balancing and CBPS"
author: "Eric Parajon with code adapted from Colin Case"
output: html_document
header-include:
  - \usepackage{amsbsy}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \DeclareMathOperator*{\argmin}{argmin} % thin space, limits underneath in displays
---

@@

Randomized controlled trials (RCTs) are often seen as the ``gold standard'' for estimating the effects of treatments, interventions, and exposures on outcomes. However, in the social sciences, it is not always possible to achieve randomization (and even in experimental studies, this doesn't happen either!), and we may have unobserved confounding variables that affect someone's likelihood of receiving treatment. Weighting helps adjust for the covariate imbalance between the treated and the control groups in a study. The weights on observations can help us "balance-out" the influence of each observation on the causal relationship. In this manner, we are making it so the treatment and control groups are *balanced* on average when considering observable characteristics (we obviously cannot balance on unobservable characteristics).

@@

stargazer(_____, _____, _____, type="html",
          covariate.labels = c("Primed",
                               "Political Interest",
                               "Female",
                               "Østlandet",
                               "Sørlandet",
                               "Vestlandet",
                               "Trøndelag",
                               "Nord-Norge",
                               "Completed Upper Secondary Education",
                               "Completed University/ College Education",
                               "Born between 1940-1949",
                               "Born between 1950-1959",
                               "Born between 1960-1969",
                               "Born between 1970-1979",
                               "Born between 1980-1989",
                               "Born in or after 1990",
                               "Intercept"),
          se=list(NULL, NULL, corr_se),
          column.labels = c("No Weights OLS", "PS Weights SWLS", "PS/NCP Weights SWLS"),
          star.cutoffs = c(.10,.05,.01,.001),
          star.char = c("+","*","**","***"),
          column.sep.width = "10pt",
          notes = c("p<.10+, p<.05*, p<.01**, p<.001***"),
          single.row = TRUE,
          notes.append=FALSE)
```

From the three models reported, we see that there is a small positive effect of being primed the cost of the existing municipal structure. Across three models, the effect is significant, p<.10+. These primed respondents are likely to increase their rating of the conservative party .1 point more on a 1-7 scale. The effect size is modest but it is statistically significant in all three models, regardless of the weighting method used.

>>>>

---
title: "Lab 12: Covariate Balancing and CBPS"
author: "Eric Parajon with code adapted from Colin Case"
output: html_document
header-include:
  - \usepackage{amsbsy}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \DeclareMathOperator*{\argmin}{argmin} % thin space, limits underneath in displays
---

@@

Randomized controlled trials (RCTs) are often seen as the "gold standard" for estimating the effects of treatments, interventions, and exposures on outcomes. However, in the social sciences, it is not always possible to achieve randomization (and even in experimental studies, this doesn't happen either!), and we may have unobserved confounding variables that affect someone's likelihood of receiving treatment. Weighting helps adjust for the covariate imbalance between the treated and the control groups in a study. The weights on observations can help us "balance-out" the influence of each observation on the causal relationship. In this manner, we are making it so the treatment and control groups are *balanced* on average when considering observable characteristics (we obviously cannot balance on unobservable characteristics).

@@

stargazer(_____, _____, _____, type="html",
          covariate.labels = c("Primed",
                               "Political Interest",
                               "Female",
                               "Østlandet",
                               "Sørlandet",
                               "Vestlandet",
                               "Trøndelag",
                               "Nord-Norge",
                               "Completed Upper Secondary Education",
                               "Completed University/ College Education",
                               "Born between 1940-1949",
                               "Born between 1950-1959",
                               "Born between 1960-1969",
                               "Born between 1970-1979",
                               "Born between 1980-1989",
                               "Born in or after 1990",
                               "Intercept"),
          se=list(NULL, NULL, corr_se),
          column.labels = c("No Weights", "PS Weights", "PS/NCP Weights"),
          star.cutoffs = c(.10,.05,.01,.001),
          star.char = c("+","*","**","***"),
          column.sep.width = "10pt",
          notes = c("p<.10+, p<.05*, p<.01**, p<.001***"),
          single.row = TRUE,
          notes.append=FALSE)
```
